---
title: "IS607 Week 8 Assignment"
author: "KarenWeigandt"
date: "October 17, 2015"
output: html_document
---

Create 3 data files, storing the same information in 3 different formats : an html table, xml and json.
Load data from the 3 files  into separate R dataframes.
Are the three dataframes identical?

First we load packages that we will be using
```{r}
library(XML)
library(plyr)
library(jsonlite)
```


we will start with the html table
```{r}
html_info_source <- "https://raw.githubusercontent.com/karenweigandt/IS607/master/FairyTales.html" ## Get the source file
html_info <- readLines(con = html_info_source) ## Read the info into a variable
html_info_2 <- readHTMLTable(html_info, stringsAsFactors = FALSE) ## extract the table info
html_info_df <- as.data.frame(html_info_2) ## convert list to data frame 
html_info_df
```

Next the xml file
```{r}
xml_info_source <- "https://raw.githubusercontent.com/karenweigandt/IS607/master/FairyTales.xml" ## Get the source file
xml_info <- readLines(con = xml_info_source) ## Read the info into a variable
xml_info_2 <- xmlToList(xml_info) ## Read the info into a list variable to maintain the list of 2 author names
xml_info_df <- ldply(xmlToList(xml_info), data.frame) ## convert list to data frame
xml_info_df
```

And last, the json file
```{r}
json_info_source <- fromJSON("https://raw.githubusercontent.com/karenweigandt/IS607/master/FairyTales.json") ## Get the source file
json_info_df <- data.frame(json_info_source, stringsAsFactors = FALSE)  ## make a dataframe
json_info_df
```

One thing I noticed with json is that the package you try makes a huge difference. 
I am also not quite sure why for the html and xml files it tells me my final line is incomplete.

The 3 data frames are definitely not identical.